%% TODO: Fix formatting here too :-)
\newcommand{\perk}[2]{\belowpdfbookmark{#1}{#1}\item[#1] (Cost: #2 \emph{CCP})}

\chapter{Perks}
\label{chap:Perks}

\section{Overview}
\label{sec:5-Overview}

Perks are special tricks, feats or other particularities that your character
is capable of doing. They can be taken once per character (unless otherwise
noted) and usually come with requirements that must be met, before they may
be bought with \emph{CPP}.

You can buy a perk anytime you like, as long as you meet the requirements.
Some perks are more powerful than others, and you may wish to consult with
your \emph{GM} first before taking perks.

\section{How Perks Work}
\label{sec:5-How-Perks-Work}

The bonuses given by perks stack with other bonuses from your character
(for example skills) but do not stack with other perks. So if one perk gives
you a +2 on small arms when firing a handgun, and another gives a +1 bonus on
small arms regardless of which is used, the +2 and +1 do not stack to +3 when
using hand guns. Some perks deliberately stack with others, but if that is the
case it is noted in the description of the perk.

\section{Perks}
\label{sec:5-Perks}

\begin{description}
  \perk{Marksman}{10} Requires \emph{Small Arms} rank 5 and \emph{Dexterity}
  at rank 7. If you are in the prone firing position you gain double the
  benefit (-6 on the difficulty instead of -3).

  \perk{Quick Draw}{10} Requires \emph{Dexterity} 7. You can draw a
  pistol or a knife that is in an slot of your vest as a free action
  in your turn. You can also draw a knife or a pistol hidden on your
  body (but still in a holster) as a free action on your turn.

  \perk{Surgeon}{5} Requires \emph{Medicine} rank 3. You are a trained
  surgeon and gain a +3 bonus on all \emph{Medicine} checks involving
  surgery. You can also perform hour long surgeries (up to 12 hours)
  without suffering in your checks or require rest.

  \perk{Toughness I}{3} Requires \emph{Constitution} 5. You immediately gain
  3 \emph{hit points} and a +1 bonus to every saving throw involving
  \emph{Constitution} to resist poison, disease or falling unconscious.

  \perk{Toughness II}{5} Requires \emph{Constitution} 6 and \emph{Toughness
    I}. You immediately gain 3 \emph{hit points} and a +1 bonus to every
  saving throw involving \emph{Constitution} to resist poison, disease or
  falling unconscious. These bonuses stack with those of \emph{Toughness I}.

  \perk{Toughness III}{7} Requires \emph{Constitution} 7 and \emph{Toughness
    II}. You immediately gain 3 \emph{hit points} and a +1 bonus to every
  saving throw involving \emph{Constitution} to resist poison, disease or
  falling unconscious. These bonuses stack with those of \emph{Toughness I} and
  \emph{Toughness II}.

  \perk{Trusty Sidearm}{5} Requires \emph{Small Arms} rank 2. With the
  \emph{Trusty Sidearm} perk you get a +2 on all checks involving your
  sidearm (including aiming, hiding it etc.).

\end{description}
